Retrofitting Effect Handlers onto OCaml

### 2020-11-05 retro-concurrency.tex

\begin{table}
\caption{Micro benchmarks without effects. Each entry shows percentage
	difference for Multicore OCaml over stock OCaml.}
\vspace{-5mm}
{\small
\begin{tabular}{r c c c c c c c c c c}
	& \rot{exnval} & \rot{exnraise} & \rot{extcall} & \rot{callback} & \rot{ack}
	& \rot{evenodd} & \rot{fib} & \rot{motzkin} & \rot{sudan} & \rot{tak} \\ \hline
	\textbf{Time} & +0.0 & -1.9 & +17 & +65  & +5.3
								& -24 & +2.2 & +10 & +0.0 & +4.2 \\
	\textbf{Instr} & +0.0 & +0.0 & +10 & +72 & +16
								 & +0.0 & +24 & +16 & +14 & +17 \\ \hline
\end{tabular}
}
\label{tab:micro}
\end{table}

We also present micro benchmarks results in Table~\ref{tab:micro}. Since micro
benchmarks magnify micro-architectural optimisations, we also report the number
of instructions executed along with time. |exnval| perform 100 million
iterations of installing exception handlers and returning with value.
|exnraise| is similar, but raises an exception in each iteration. |extcall| and
|callback| perform 100 million external calls and callbacks to identity
functions. The other micro benchmarks are highly recursive programs and were
taken from ~\ref{Folklore}.

The results show that exceptions are no more expensive in |MC| compared to
|stock|. In the other programs, |MC| executes more instructions due to stack
overflow checks, except |evenodd| on which the check is eliminated by the
compiler. Despite this, observe that |MC| is 24\% faster than stock on
|evenodd|. At these scales, we observed that padding tightloops with a few
|nop| instructions, which changes the loop alignment, makes the code upto 15\%
faster. Hence, the difference in running times under 15\% is not statistically
significant. External calls and callbacks are more expensive due to additional
bookkeeping. Callbacks are significantly more expensive, but are also less of a
concern compared to external calls which are far more numerous. As seen
previously, these results however do not affect macro benchmark performance.

>>>>

\begin{table}
\caption{Micro benchmarks without effects. Each entry shows percentage
	difference for Multicore OCaml over stock OCaml.}
\vspace{-5mm}
{
\begin{tabular}{r c c c c c c c c c c}
	& \rot{exnval} & \rot{exnraise} & \rot{extcall} & \rot{callback} & \rot{ack}
	& \rot{fib} & \rot{motzkin} & \rot{sudan} & \rot{tak} \\ \hline
	\textbf{Time} & +0.0 & -1.9 & +17 & +65  & +5.3
								& +2.2 & +10 & +0.0 & +4.2 \\
	\textbf{Instr} & +0.0 & +0.0 & +10 & +72 & +16
								 & +24 & +16 & +14 & +17 \\ \hline
\end{tabular}
}
\label{tab:micro}
\end{table}

We also present micro benchmarks results in Table~\ref{tab:micro}. Since micro
benchmarks magnify micro-architectural optimisations, we also report the number
of instructions executed along with time. |exnval| perform 100 million
iterations of installing exception handlers and returning with value.
|exnraise| is similar, but raises an exception in each iteration. |extcall| and
|callback| perform 100 million external calls and callbacks to identity
functions. The other micro benchmarks are highly recursive programs and were
taken from ~\ref{Folklore}. For micro benchmarks, we observed that padding
tightloops with a few |nop| instructions, which changes the loop alignment,
makes the code upto 15\% faster. Hence, the difference in running times under
15\% may not be statistically significant.

The results show that exceptions are no more expensive in |MC| compared to
|stock|. In the other programs, |MC| executes more instructions due to stack
overflow checks. External calls and callbacks are more expensive due to
additional bookkeeping. Callbacks are significantly more expensive in |MC|.
Since we reuse the current fiber stack for callbacks, we need to ensure it has
enough room for inserting additional frames, while |stock| does not need to do
this. Callback performance is less important than external calls, which are far
more numerous. As seen previously, these results however do not affect macro
benchmark performance.
